Model Adaptivity for Goal-Oriented Inference using Adjoints

### 2016-11-15 paper/algorithms_v3.tex
stashing edits
removed linearization about psi in superadj as a source of error

Just as error estimates can be used to guide mesh refinement~\cite{BecRann01,VendDarm00,PrudOden99,BecVex05,AinsOden11}, the error estimate \cref{eq:finErrExp} can be localized to give elemental contributions and used to guide physics-based refinement over the domain to create a mixed-fidelity model. The error estimate can be calculated again, using the mixed-fidelity model as the lower-fidelity model. This process can be repeated, successively increasing the proportion of the domain in which the high-fidelity model is used, until some threshold is reached. \Cref{alg:refSeries} describes this approach. Note that the QoI error adjoint problem \cref{eq:superAdjEq} involves linearization about $\Psi_{HF}$, which is not available, so in the case of a nonlinear goal-oriented inverse problem, the QoI error adjoint problem is approximated by linearizing about $\Psi_{LF}$ instead.
%
\alglanguage{pseudocode}
\begin{algorithm}[h!]
\small
\caption{An algorithm to adaptively build a mixed-fidelity model for low error in the QoI.}
\label{alg:refSeries}
\begin{algorithmic}[1]
\State{Define maximum acceptable absolute relative QoI error \texttt{errTol}}
\State{Define maximum number of adaptive iterations \texttt{maxIter}}
\Procedure{$\texttt{BuildMixedFidelity}$}{HF model, LF model, \texttt{errTol}, \texttt{maxIter}}
	\State{Let the model MF$_0$ be the LF model applied everywhere in the domain.}
	\State{$i\gets0$}
	\State{Solve for stationary point $\Psi_{MF_0}$ of augmented Lagrangian $\mathcal{M}_{MF_0}$}
	\State{Solve QoI error adjoint equation, linearized about $\Psi_{MF_0}$, for \par\hskip\algorithmicindent supplementary adjoint $\Lambda_0$ (see \cref{eq:superAdjEq})}
	\State{Compute QoI error estimate

	$\epsilon_0=-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{MF_0})(\Lambda_0)+\mathcal M_{HF}(\Psi_{MF_0})-\mathcal M_{MF_0}(\Psi_{MF_0})$}
	\State{Calculate QoI $I(q_{MF_0},u_{MF_0})$}
	\While{$i<$ \texttt{maxIter} and $|\epsilon_i/(\epsilon_i+I(q_{MF_i},u_{MF_i}))|>$ \texttt{errTol}}
		\State{\begin{varwidth}[t]{\linewidth}Localize $\epsilon_i$ (see \cref{sec:errLocal}) and use this decomposition to guide \par\hskip\algorithmicindent formation of new mixed-fidelity model MF$_{i+1}$\end{varwidth}}
		\State{$i\gets i+1$}
		\State{Solve for stationary point $\Psi_{MF_i}$ of augmented Lagrangian $\mathcal{M}_{MF_i}$}
		\State{Solve QoI error adjoint equation, linearized about $\Psi_{MF_i}$, for

@@

%------------------------------------------------------------------------------------------------------------------------%
%\subsection{Extension to Offline-Online Setting}\label{sec:onoff}
%------------------------------------------------------------------------------------------------------------------------%

%\Cref{alg:refSeries} can be naturally extended to an offline-online setting. One variation would be analagous to the offline-online decomposition proposed in \cite{LiebWill13}. The offline phase would consist of adaptively creating a mixed-fidelity model with an appropriate error tolerance given some expected observations $d^*$. When new data is received, one can then solve the inverse problem with the chosen mixed-fidelity model and the new data, and, if desired, compute an error estimate for the QoI. The mixed-fidelity inverse problems can be expected to require less time to solve than the high-fidelity inverse problems.

%Another variation occurs in the context of a quasi-incremental data assimilation approach. A multi-fidelity model adaptive constraint can be generated for a given dataset $d_1$ in the offline phase. In the online phase, additional data points are included to form an augmented dataset $d_2$, and instead of solving the complete inverse problem with all the data points and the high-fidelity constraints, an iterative algorithm can be initiated using the multi-fidelity representation developed in the offline phase. The adaptive procedure, with the supplementary adjoint can be used to adapt the preexisting multi-fidelity representation to the new data points. In such a situation where $d_1\in d_2$, we expect that the model refinement will be concentrated around those of the new data points that also inform the QoI. Thus, in the online phase, even though the inverse problem will be solved with all data points $d_2$, the use of high-fidelity models will be limited, and the bulk of the computation (computing the auxilary variables, and supplementary adjoint) will be focused on identifying those new data points, which have the maximum impact on the QoI.

>>>>

Just as error estimates can be used to guide mesh refinement~\cite{BecRann01,VendDarm00,PrudOden99,BecVex05,AinsOden11}, the error estimate \cref{eq:finErrExp} can be localized to give elemental contributions and used to guide physics-based refinement over the domain to create a mixed-fidelity model. The error estimate can be calculated again, using the mixed-fidelity model as the lower-fidelity model. This process can be repeated, successively increasing the proportion of the domain in which the high-fidelity model is used, until some threshold is reached. \Cref{alg:refSeries} describes this approach. 
%Note that the QoI error adjoint problem \cref{eq:superAdjEq} involves linearization about $\Psi_{HF}$, which is not available, so in the case of a nonlinear goal-oriented inverse problem, the QoI error adjoint problem is approximated by linearizing about $\Psi_{LF}$ instead.
%
\alglanguage{pseudocode}
\begin{algorithm}[h!]
\small
\caption{An algorithm to adaptively build a mixed-fidelity model for low error in the QoI.}
\label{alg:refSeries}
\begin{algorithmic}[1]
\State{Define maximum acceptable absolute relative QoI error \texttt{errTol}}
\State{Define maximum number of adaptive iterations \texttt{maxIter}}
\Procedure{$\texttt{BuildMixedFidelity}$}{HF model, LF model, \texttt{errTol}, \texttt{maxIter}}
	\State{Let the model MF$_0$ be the LF model applied everywhere in the domain.}
	\State{$i\gets0$}
	\State{Solve for stationary point $\Psi_{MF_0}$ of augmented Lagrangian $\mathcal{M}_{MF_0}$}
	\State{Solve QoI error adjoint equation
	%, linearized about $\Psi_{MF_0}$,
	 for \par\hskip\algorithmicindent supplementary adjoint $\Lambda_0$ (see \cref{eq:superAdjEq})}
	\State{Compute QoI error estimate

	$\epsilon_0=-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{MF_0})(\Lambda_0)+\mathcal M_{HF}(\Psi_{MF_0})-\mathcal M_{MF_0}(\Psi_{MF_0})$}
	\State{Calculate QoI $I(q_{MF_0},u_{MF_0})$}
	\While{$i<$ \texttt{maxIter} and $|\epsilon_i/(\epsilon_i+I(q_{MF_i},u_{MF_i}))|>$ \texttt{errTol}}
		\State{\begin{varwidth}[t]{\linewidth}Localize $\epsilon_i$ (see \cref{sec:errLocal}) and use this decomposition to guide \par\hskip\algorithmicindent formation of new mixed-fidelity model MF$_{i+1}$\end{varwidth}}
		\State{$i\gets i+1$}
		\State{Solve for stationary point $\Psi_{MF_i}$ of augmented Lagrangian $\mathcal{M}_{MF_i}$}
		\State{Solve QoI error adjoint equation%, linearized about $\Psi_{MF_i}$,
		 for

@@

### 2016-11-15 paper/numerical_experiments_v3.tex
removed linearization about psi in superadj as a source of error

\Cref{fig:baseRef} shows the local error contributions, as well as the subdomains where the low- and high-fidelity models are used, for the series of mixed-fidelity models thus generated. Each linear Lagrange basis function's contribution is plotted at its nonzero node.
%
\begin{figure}[htbp]
%\captionsetup[subfloat]{captionskip=-5pt}
\centering
\subfloat[LF $\equiv$ MF$_0$ ($0\%$ HF)]{
  \includegraphics[width=0.46\textwidth]{baseSeries/cd_cdr_LF_divvy.png}
  \includegraphics[width=0.49\textwidth]{baseSeries/err_breakdown_LF.png}
  \label{fig:baseRef0}
} \\
\subfloat[MF$_1$ ($5\%$ HF)]{
  \includegraphics[width=0.46\textwidth]{baseSeries/cd_cdr_MF01_divvy.png}
  \includegraphics[width=0.49\textwidth]{baseSeries/err_breakdown_MF01.png}
} \\
\subfloat[MF$_2$ ($10\%$ HF)]{
  \includegraphics[width=0.46\textwidth]{baseSeries/cd_cdr_MF02_divvy.png}
  \includegraphics[width=0.49\textwidth]{baseSeries/err_breakdown_MF02.png}
} \\
\caption{Left: Multi-fidelity refinement over the domain (low-fidelity convection-diffusion model used in red portion, high-fidelity convection-diffusion-reaction model used in blue portion). Right: local error contributions. }
\label{fig:baseRef}
\end{figure}
%
Note that the error contribution of each basis function whose support is entirely within the high-fidelity regions is zero.

@@

We notice that, compared to the results in \cref{fig:baseErr}, the error estimates for the low-fidelity and first mixed-fidelity models are far from the true errors. This can be attributed to the linearization about $\Psi_{LF}$ and $\Psi_{MF_{1}}$ instead of $\Psi_{HF}$ in solving the supplementary adjoints; the large differences in the QoI for the LF and MF01 models compared to the high-fidelity QoI indicate that $\Psi_{LF}$ and $\Psi_{MF_{1}}$ are significantly different from $\Psi_{HF}$. Compared to the pair of models in \cref{sec:cdvcdrSetup}, the low- and high-fidelity models in this case are more dissimilar, even though in both \cref{sec:cdvcdrSetup} and this case the nonlinear term in the high-fidelity model \cref{eq:cdvcdrHF} and \cref{eq:cdvcdrHF3D} is a quadratic reaction term. % KW: don't know what last part of this sentence means. what hi-fi models?
The possibility of large inaccuracies in the error estimates of low-fidelity and mostly unrefined mixed-fidelity models suggests that one should use more than just the error estimate to decide when to stop the adaptive algorithm; a possible additional criterion could be to require that the estimated high-fidelity QoI ($I(q_{MF},u_{MF})+\epsilon_i$) for two consecutive iterations agree to within some tolerance. % KW: what behavior? this last sentence doesn't follow from the previous one. and what predicted hi-fi QOI? Do you mean mixed-fi? I'm not sure this is a good idea. I'd probably delete this last sentence.
% or just let the reader decide how to deal with possibly inaccurate initial error estimates?

>>>>

\Cref{fig:baseRef} shows the local error contributions, as well as the subdomains where the low- and high-fidelity models are used, for the series of mixed-fidelity models thus generated. %Each linear Lagrange basis function's contribution is plotted at its nonzero node.
%
\begin{figure}[htbp]
%\captionsetup[subfloat]{captionskip=-5pt}
\centering
\subfloat[LF $\equiv$ MF$_0$ ($0\%$ HF)]{
  \includegraphics[width=0.46\textwidth]{baseSeries/cd_cdr_LF_divvy.png}
  \includegraphics[width=0.49\textwidth]{baseSeries/err_breakdown_LF.png}
  \label{fig:baseRef0}
} \\
\subfloat[MF$_1$ ($5\%$ HF)]{
  \includegraphics[width=0.46\textwidth]{baseSeries/cd_cdr_MF01_divvy.png}
  \includegraphics[width=0.49\textwidth]{baseSeries/err_breakdown_MF01.png}
} \\
\subfloat[MF$_2$ ($10\%$ HF)]{
  \includegraphics[width=0.46\textwidth]{baseSeries/cd_cdr_MF02_divvy.png}
  \includegraphics[width=0.49\textwidth]{baseSeries/err_breakdown_MF02.png}
} \\
\caption{Left: Multi-fidelity refinement over the domain (low-fidelity convection-diffusion model used in red portion, high-fidelity convection-diffusion-reaction model used in blue portion). Right: local error contributions. }
\label{fig:baseRef}
\end{figure}
%
Note that the error contribution of each basis function whose support is entirely within the high-fidelity regions is zero.

@@

We notice that, compared to the results in \cref{fig:baseErr}, the error estimates for the low-fidelity and first mixed-fidelity models are far from the true errors. This can be attributed to the 
%linearization about $\Psi_{LF}$ and $\Psi_{MF_{1}}$ instead of $\Psi_{HF}$ in solving the supplementary adjoints; 
third-order term that is ignored in the error estimate;
the large differences in the QoI for the LF and MF01 models compared to the high-fidelity QoI indicate that $\Psi_{LF}$ and $\Psi_{MF_{1}}$ are significantly different from $\Psi_{HF}$. Compared to the pair of models in \cref{sec:cdvcdrSetup}, the low- and high-fidelity models in this case are more dissimilar, even though in both \cref{sec:cdvcdrSetup} and this case the nonlinear term in the high-fidelity model \cref{eq:cdvcdrHF} and \cref{eq:cdvcdrHF3D} is a quadratic reaction term. % KW: don't know what last part of this sentence means. what hi-fi models?
The possibility of large inaccuracies in the error estimates of low-fidelity and mostly unrefined mixed-fidelity models suggests that one should use more than just the error estimate to decide when to stop the adaptive algorithm; a possible additional criterion could be to require that the estimated high-fidelity QoI ($I(q_{MF},u_{MF})+\epsilon_i$) for two consecutive iterations agree to within some tolerance. % KW: what behavior? this last sentence doesn't follow from the previous one. and what predicted hi-fi QOI? Do you mean mixed-fi? I'm not sure this is a good idea. I'd probably delete this last sentence.
% or just let the reader decide how to deal with possibly inaccurate initial error estimates?
